PeerAssignment 1 
========================================================
 1) set the working directory 
 2) Read the i/p file into a file named activity
 3) Remove all "NA"
 4) split the source file based on dates using tapply function
 5) Find the number of levels ie 
 6) Find the total number of steps for each of the date  using for loop
 7) Display the total steps taken per day as a histogram using ggplot2

## Loading and preprocessing the data

 1) Code for Loading the data

```{r}
 setwd("C:/Users/venkat rao/Documents/Coursera/RR")
 activity<-read.csv("activity.csv",stringsAsFactors=FALSE)

```

 2)Code for preprocessing: Removing NA for analysis

```{r}
  histactivity<-na.omit(activity)
```
## What is mean total number of steps taken per day?

 Applying tapply to find total steps per day 

```{r}
    splitsum<-tapply(histactivity$steps,histactivity$date,sum,simplify=FALSE) 
    histactivity$date<-as.factor(histactivity$date)
    index<-length(levels(histactivity$date))
 

    for ( i in 1:index )
         {    if ( i==1)
             {
                  steps<-splitsum[[i]]
                  date<-dimnames(splitsum)[[1]] [i]  
                  Totalstepsperday<-data.frame(date,steps)   }
      else
       
             {   steps<-splitsum[[i]]
                 date<-dimnames(splitsum)[[1]] [i]  
                 Totalstepsperday1<-data.frame( date,steps)
                 Totalstepsperday<-rbind(Totalstepsperday,Totalstepsperday1)
              }
            }   
        
  library(ggplot2)
  library(scales) 
```
  1) Make a histogram of the total number of steps taken each day

  Histogram of Total no:of steps vs Date

```{r}
     Totalstepsperday$date<-as.Date(Totalstepsperday$date)
    m <- ggplot(Totalstepsperday, aes(x=date))
    m + geom_histogram(aes(weight = steps)) + scale_x_date(breaks= "1 weeks",minor_breaks = "2 day",labels = date_format("%m/%d")) 
```
  2) Calculate and report the mean and median total number of steps taken per day 

```{r Calculate median mode ,echo=FALSE}
     options(scipen=1,disgits=2)   
     mean<-mean(Totalstepsperday$steps)
     median<-median(Totalstepsperday$steps) 
     mean<-trunc(mean)  
     
       
```
     Median number of steps is  `r median` & Mean number of steps  is   `r mean`.
     
## What is the average daily activity pattern?
     
 1) Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged  across all days (y-axis)
    1a) Apply tapply to find  average number of steps across each interval
    1b) Create a dataframe Averagesteps 
    
```{r average daily activity pattern }
 
    splitinterval<-tapply(histactivity$steps,histactivity$interval,mean,simplify=FALSE) 
    histactivity1<-histactivity
    histactivity1$interval<-as.factor(histactivity1$interval)
    index<-length(levels(histactivity1$interval))
    for ( i in 1:index )
    {     if ( i==1)
              {
                  steps<-splitinterval[[i]]
                  interval<-as.numeric(dimnames(splitinterval)[[1]] [i])  
                  Averagesteps<-data.frame(interval,steps)   
              }
         else
            {      steps<-splitinterval[[i]]
                   interval<-as.numeric(dimnames(splitinterval)[[1]] [i])  
                   Averagesteps1<-data.frame(interval,steps)
                   Averagesteps<-rbind(Averagesteps,Averagesteps1)
            }
    }        
         
    plot(Averagesteps$interval,Averagesteps$steps,type="l",xlab="Interval",ylab="Number of Steps",xlim=c(0,2400))
```
    
    
  
2) Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r Calculate Max steps ,echo=FALSE}
         
       Averagesteps<-Averagesteps[order(-(Averagesteps$steps)),]
       maxstepsinterval<-Averagesteps[1,1]
       missingvalue<-apply(is.na(activity),2,sum)  
       
```
      The 5 minute interval having Avergae maimum number of steps is  `r maxstepsinterval`.
##  Imputing missing values

   1) The total number of missing values in the dataset (i.e. the total number of rows with NAs) is `r missingvalue `

   2) Filled the missing values with mean of steps across each interval.  
   
   3) Dataset activityna created which is filled with missing values .These values are the mean of the days per interval..

   
   
```{r histogram ,echo=FALSE}
     options(scipen=1,disgits=2)   
     library(gam)
     histactivity1<-histactivity
     histactivity1$interval<-as.factor(histactivity1$interval)
     index<-length(levels(histactivity1$interval))
     naintervals<-split(activity,activity$interval,drop=TRUE)
     for ( i in 1:index )
         {    
          activityna1<-as.data.frame(naintervals[i])
          names(activityna1)[1]="steps" 
          names(activityna1)[2]="date" 
          names(activityna1)[3]="interval" 
          if (i ==1 ) {      activityna<-na.gam.replace(activityna1)      }
          else
                      {      activityna1<-na.gam.replace(activityna1)  
                             activityna<-rbind(activityna1,activityna)    }
         } 
      activityna<-activityna[order(activityna$date),]
              splitsumna<-tapply(activityna$steps,activityna$date,sum,simplify=FALSE) 
              hist<-activity
              hist$date<-as.factor(hist$date)
              index<-length(levels(hist$date))
 

    for ( i in 1:index )
         {    if ( i==1)
             {
                  steps<-splitsumna[[i]]
                  date<-dimnames(splitsumna)[[1]] [i]  
                  Totalstepsperdayna<-data.frame(date,steps)   }
      else
       
             {   steps<-splitsumna[[i]]
                 date<-dimnames(splitsumna)[[1]] [i]  
                 Totalstepsperdayna1<-data.frame( date,steps)
                 Totalstepsperdayna<-rbind(Totalstepsperdayna,Totalstepsperdayna1)
              }
            }   
       library(ggplot2)
       library(scales)  
     Totalstepsperdayna$date<-as.Date(Totalstepsperdayna$date)
     m <- ggplot(Totalstepsperdayna, aes(x=date))
     m + geom_histogram(aes(weight = steps)) + scale_x_date(breaks= "1 weeks",minor_breaks = "2 day",labels = date_format("%m/%d"))
       
```

```{r Calculate new median & mode ,echo=FALSE}
     options(scipen=1,disgits=2)   
     mean<-mean(Totalstepsperdayna$steps)
     median<-median(Totalstepsperdayna$steps) 
     mean<-trunc(mean)  
            
``` 
The mean and median total number of steps taken per day is  `r mean ` & `r median `

Values dont differ much from the estimates from the first part of the assignment?  

Impact according to my computations is minimal

## Are there differences in activity patterns between weekdays and weekends?

        Based on the plots below we can see there is not much of a difference between both the plots

1) Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
    Created ..

2) Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r time series plot  ,echo=FALSE}
       
       df2<-activityna
       df2$day<-weekdays(as.Date(Totalstepsperdayna$date))
       df2<-within(df2,{day[day=="Monday" | day=="Tuesday" | day=="Wednesday" | day=="Thursday" | day=="Friday"]<-"Weekday"})
       df2<-within(df2,{day[day=="Saturday" | day=="Sunday"]<-"Weekend"})
               
       daysplit<-split(df2,df2$day)
       weekday<-as.data.frame(daysplit[1]) 
       names(weekday)[1]="steps" 
       names(weekday)[2]="date" 
       names(weekday)[3]="interval"
       names(weekday)[4]="day" 
 
       weekdaysplit<-tapply(weekday$steps,weekday$interval,mean,simplify=FALSE) 
       histactivity1<-weekday
       histactivity1$interval<-as.factor(histactivity1$interval)
       index<-length(levels(histactivity1$interval))


       for ( i in 1:index )
            {   if ( i==1)
                   {
                        steps<-weekdaysplit[[i]]
                        interval<-as.numeric(dimnames(weekdaysplit)[[1]] [i])  
                        day ="weekdays"
                        Averagedaysteps<-data.frame(interval,steps,day)   
                    }
                 else
                    {    steps<-weekdaysplit[[i]]
                         interval<-as.numeric(dimnames(weekdaysplit)[[1]] [i])  
                         day ="weekdays"
                         Averagedaysteps1<-data.frame(interval,steps,day)
                         Averagedaysteps<-rbind(Averagedaysteps1,Averagedaysteps)
                      }
              }   

          weekend<-as.data.frame(daysplit[2])  
          names(weekend)[1]="steps" 
          names(weekend)[2]="date" 
          names(weekend)[3]="interval"
          names(weekend)[4]="day"  
          weekendsplit<-tapply(histactivity$steps,histactivity$interval,mean,simplify=FALSE) 
          histactivity1<-weekend
          histactivity1$interval<-as.factor(histactivity1$interval)
          index<-length(levels(histactivity1$interval))


        for ( i in 1:index )
            {     steps<-weekendsplit[[i]]
                  interval<-as.numeric(dimnames(weekendsplit)[[1]] [i])  
                  day ="weekend"
                  Averagedaysteps1<-data.frame(interval,steps,day)
                  Averagedaysteps<-rbind(Averagedaysteps,Averagedaysteps1)
            }
              

       library(lattice)
       xyplot(steps~interval | day ,type = c("l"),layout=c(1,2),data=Averagedaysteps)        
       
        
``` 
     